---
title: "Analytics_III_Problem_I"
author: "Shubham Shinde"
date: "`r Sys.Date()`"
output:
  html_document: default
  pdf_document: default
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(fig.width=10, fig.height=7) # Fixed the margin
library(ggplot2)
library(dplyr)
library(tidyverse)
library(rpart)
library(MASS)
library(klaR)
```

# Table of Contents
1. [Outline](#outline)
2. [Miles Per Gallon Analysis ](#mpg)
3. [1. Data Import](#imp)
4. [2. Data Cleaning](#cleaning)
5. [3. Descriptive Data Analysis](#desc)
6. [4.  Plots](#plots)
7. [5. Model](#models)


# Outline
- Decide, if you want to use Markdown or Sweave/knit.
- Analyze the mpg data set [link](https://archive.ics.uci.edu/ml/datasets/auto+mpg) with one statistical method of your choice.
- Prepare a reproduce-able document of your analysis with Markdown or Sweave/knit.
- Alternative: choose another data set.

# Miles Per Gallon Analysis 

## 1. Data Import
```{r dataimport, }
data <- read.csv("C:/Users/shubh/OneDrive - SRH IT/University/Semester 3/Analytics 3/Codes/auto-mpg.data",na = "?")
head(data)
```



## 2. Data Cleaning
```{r datacleaning, }
data$horsepower <- gsub("?",NA,data$horsepower, fixed = TRUE)
s <- colnames(data)
data <-  data %>%
  drop_na()
```

```{r datatrans, }
data$mpg <- as.numeric(data$mpg)
data$horsepower <- as.numeric(data$horsepower)
# data$cylinders <- as.factor(data$cylinders)
data$car.name <- as.factor(data$car.name)
# data$model.year <- as.factor(data$model.year)
```



```{r datatrans1, }
data <- data %>% 
  mutate(origin_alpha = as.factor(case_when(
    origin == 1 ~ "American",
    origin == 2 ~ "European",
    origin == 3 ~ "Japanese",
    )))
```
## 3. Descriptive Data Analysis

The summary of the dataset is as follows:
```{r datasumm, }
summary(data)
```
We can check some data examples:
```{r datahead, }
head(data)
```
Lets check the structure of the data:
```{r datastr, }
str(data)
```


## 4.  Plots


### MPG Vs CYL


```{r mpgvscyl}
ggplot(data = data, aes(cylinders, mpg)) + geom_boxplot(aes(color = cylinders))
```
Higher cyl has lower mpg

### MPG Vs displacement
```{r mpgvsdisp}
ggplot(data = data, aes(displacement, mpg)) + geom_jitter()
```
Higher Displacement lower mpg

### MPG Vs horsepower
```{r mpgvshor}
ggplot(data = data, aes(horsepower, mpg)) + geom_jitter()
```
Higher Horsepower lower mpg

### MPG vs weight
```{r mpgvswe}
ggplot(data = data, aes(weight, mpg)) + geom_jitter()
```
Higher Weight lower mpg

### MPG vs acceleration

```{r mpgvsacc}
ggplot(data = data, aes(acceleration, mpg)) + geom_point()
```
MPG increases with acceleration
Average Acceleration has better mpg

### MPG vs Model Year
```{r mpgvsmod}
ggplot(data = data, aes(x = model.year, y = mpg)) + geom_boxplot()
```
Recent cars are more efficient, have better mpg.


## Correlation Matrix - Filter Factorial/Char Columns from Dataframe
```{r corfilter}
exclude_vec <- c("model.year", "car.name","cylinders")
cor_data <- data %>% 
  dplyr::select(setdiff(s, exclude_vec))
```

## Correlation Matrix
```{r cor}
cor(cor_data)
```



## 5. Model
### Independent Variables 
### 1. Prediction of mpg

```{r datafil}
data_cart <- data %>%
  dplyr::select(-car.name, -origin_alpha)
```


## Decision Tree
```{r datafil1}
cart <- rpart(mpg ~ ., data = data_cart)
cart$variable.importance
```
## Pruning
```{r datafil2}

cart.p <- prune(cart, cp = .05)
cart.p$variable.importance
```

```{r datafil3}

test_data <- data.frame(weight = c(1600),model.year = c(06), acceleration = c(56),horsepower = c(100), origin = c(2), cylinders = c(4), displacement = c(134))
test_data
```
```{r datafil4}

predict(cart, newdata = test_data)
```

## Linear Regression

# '''
# Null Hypothesis: There is no relationship between independent variables and mpg
# '''

```{r datafil5}
lin.reg <- lm(mpg ~ . - origin_alpha, data = data)
summary(lin.reg)
```
```{r datafil6}
lin.reg <- lm(mpg ~ cylinders + horsepower + weight + origin, data = data)
summary(lin.reg)
```
```{r datafil7}
test_data <- data.frame(weight = c(1600),model.year = c(06), acceleration = c(56),horsepower = c(100), origin = c(2), cylinders = c(4), displacement = c(134))
test_data
```
```{r datafil8}
predict(lin.reg, newdata = test_data)
```
## Classification Problem
## What is the origin of a car


## Linear Discriminant Analysis
```{r datafil9}
data_lda <- data %>%
  dplyr::select(-origin, -car.name)
```
```{r datafil10}
lda.model <- lda(origin_alpha ~ .,data = data_lda)
```
```{r datafil11}
summary(lda.model)
```
```{r datafil12}
lda.model
```
```{r datafil13}
test_data <- data.frame(mpg = c(45),weight = c(1600),model.year = c(06), acceleration = c(56),horsepower = c(100), origin = c(2), cylinders = c(4), displacement = c(134))
```
```{r datafil14}
test_data
```
```{r datafil15}
predict(lda.model, newdata = test_data)
```

## Greedy Wilks
```{r datafil16}
fit.gw <- greedy.wilks(origin_alpha ~ . , data = data_lda)
fit.gw
```
```{r datafil17}
lda_gw <- lda(fit.gw$formula, data = data_lda)
lda_gw
```
```{r datafil18}
test_data <- data.frame(mpg = c(45), weight = c(1600),model.year = c(06), acceleration = c(56),horsepower = c(100), origin = c(2), cylinders = c(4), displacement = c(134))
test_data
```
```{r datafil19}
p <- predict(lda.model, newdata = test_data)
p
```
